Types for R, Empirically Designed and Evaluated

### 2020-05-09 evaluation/R/corpus-analysis.Rmd
Draft of 2

---
title: "Package Corpus Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  min_revdeps: 5
  min_coverage: .65
  which_dependencies:
    - Depends
    - Imports
---

@@

@@

@@

summary_corpus(pkgs, "pkgs")
```

#### Plots

@@

plot_all(pkgs_considered)
ggsave(CORPUS_PLOT)
```

@@

```{r plot corpus}
plot_corpus <- function(df, code_outlier=5e3) {
  coverage_mean <- mean(df$coverage_expr)
  coverage_median <- median(df$coverage_expr)
  revdeps_mean <- mean(df$revdeps)
  revdeps_median <- median(df$revdeps)

@@

plot_corpus(corpus)
```

### List of reverse dependencies

```{r list of reverse depenencies}
corpus_revdeps <- pkgs_revdeps_raw %>% semi_join(corpus, by="package")
write_csv(corpus_revdeps, CORPUS_REVDEPS)
write_lines(unique(corpus_revdeps$revdep), CORPUS_REVDEPS_UNIQUE)
```

>>>>

---
title: "Corpus Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  min_revdeps: 5
  min_coverage: .65
  package_size_outlier: 5000
---

@@

Sys.setenv(
  TYPER_BASE_DIR=system2(
    "git", 
    c("rev-parse", "--show-toplevel"), 
    stdout=TRUE
  )
)

@@

## Overview

### Thresholds

```{r}
overview_table(
  r("threshold code coverage", percent(params$min_coverage)),
  r("threshold revdeps", params$min_revdeps),
  r("package size outier", params$package_size_outlier)
)
```

### Table

@@

summary_corpus(pkgs, "All")
```

### Plots

@@

plot_all(pkgs_considered)
```

@@

```{r plot corpus}
plot_corpus <- function(df, code_outlier) {
  coverage_mean <- mean(df$coverage_expr)
  coverage_median <- median(df$coverage_expr)
  revdeps_mean <- mean(df$revdeps)
  revdeps_median <- median(df$revdeps)

@@

plot_corpus(corpus, params$package_size_outlier)
ggsave(CORPUS_PLOT)
```

### 2020-05-09 paper/main.tex
Draft of 2

\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\setcopyright{none}
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{my_style}
\usepackage{listings, wrapfig,xspace}

@@

\newcommand{\eg}{\emph{e.g.},\xspace}
\newcommand{\ie}{\emph{i.e.},\xspace}

@@

\newcommand{\contractr}{\emph{contractr}\xspace} % contractr
\newcommand{\roxygen}{\emph{roxygen2}\xspace} % roxygen
\newcommand{\typetracer}{\emph{typetracer}\xspace} % typetracer

@@

%
%
%
%
\subsection{The Corpus}
\label{sec:corpus}

\AT{This section needs to be tight.
Reviewers were skeptical of CRAN and using the test and example code.
We should convince them otherwise.}

\AT{Some sentences that might be useful:}
We selected packages that have both \AT{a good amount of} reverse dependencies to ensure their broad usefulness, as well \AT{sufficiently high} code and branch coverage of test, example, and vignette programs to ensure that our analysis built up an accurate picture of function use.

%
%
\subsubsection{Cross-Validation Against Kaggle}

Kaggle is a website offering an environment for hosting Jupyter notebooks, and importantly provides access to a large number of open-source data sets and R and Python code.
This solved a major issue with R programs hosted on e.g. Github, where often the data is not made available.
Our aim is to run R programs hosted on Kaggle to validate the type signatures we generated for popular R packages.

\AT{Anything else to say on Kaggle, here? We will probably talk more about the corpus we analyzed later.}

>>>>

\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\setcopyright{none}
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{my_style}
\usepackage{listings, wrapfig,xspace}
\usepackage{paralist}

@@

\newcommand{\eg}{\emph{e.g.},\xspace}
\newcommand{\ie}{\emph{i.e.},\xspace}
\newcommand{\cf}{\emph{cf.}\xspace}

@@

\newcommand{\contractr}{\emph{contractr}\xspace} % contractr
\newcommand{\roxygen}{\emph{roxygen2}\xspace} % roxygen
\newcommand{\typetracer}{\emph{typetracer}\xspace} % typetracer
\newcommand{\rdt}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}

\input{corpus.tex}

@@

%
%
%
%
\section{Project Corpus}\label{sec:corpus}

For this paper we have selected \CorpusLoadable packages consisting of
\CorpusRCodeRnd lines of R code and \CorpusNativeCodeRnd lines of native code
(C/C++/Fortran).
\FK{Add more details}
Figure~\ref{fig:corpus} shows all the packages, the size of the dots reflects
the size of the project in terms of both R and native lines of code\footnote{
Lines of source code reported excludes comments and blank lines, counted by \emph{cloc}, \cf \url{https://github.com/AlDanial/cloc}}, the x-axis
indicates the expression code coverage in percents and the y-axis gives the
number of reverse dependencies in log scale. Dotted lines indicate means.
Packages with over \PackageSizeOutierRnd lines of code are annotated.

\begin{figure}[!h]  \centering
  \includegraphics[width=.8\linewidth]{plots/corpus.pdf}
  \caption{Corpus overview}\label{fig:corpus}
\end{figure}

These packages come from the Comprehensive R Achive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \AllCranRnd packages\footnote{CRAN receives about 6 new package
  submissions a day~\cite{Ligges2017}} containing over \AllRCodeRnd and
\AllNativeCodeRnd R and native code respectively. Unlike other open code
repositories like GitHub, CRAN is a curated repository. Each submitted package
must abide to a number of well-formedness rules that are automatically checked
asserting certain quality. Most relevant for this work is that all of the
runnable code (including code snippets from examples and vignettes) is tested
and only a successfully running package is admitted in the archive.

We have downloaded and installed all available CRAN packages. Out of the
\AllCranRnd packages, we managed to install \AllLoadableRnd. The main
reason for this is that \rdt is based on R 3.5.0 and some of the packages are no
longer compatible with it. Some packages also require extra native dependencies
which were not present on our servers.
We have two criteria for including a
package into the corpus:
\begin{inparaenum}[(1)]
\item have runnable code that covers significant part of the package source
code from which type signatures could be inferred, and
\item have some reverse dependencies that allows us to evaluate the inferred
types on runnable code from these dependent packages.
\end{inparaenum}
The concrete thresholds were set to at least \ThresholdCodeCoverage of expression
coverage and at minimum \ThresholdRevdeps reverse dependencies.

The code coverage was computed for each package using
\covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R tool.
The reverse package dependencies were extracted from the package metadata using
builtin function.

% \AT{This section needs to be tight.
% Reviewers were skeptical of CRAN and using the test and example code.
% We should convince them otherwise.}

% \AT{Some sentences that might be useful:}
% We selected packages that have both \AT{a good amount of} reverse dependencies to ensure their broad usefulness, as well \AT{sufficiently high} code and branch coverage of test, example, and vignette programs to ensure that our analysis built up an accurate picture of function use.

% %
% %
% \subsubsection{Cross-Validation Against Kaggle}

% Kaggle is a website offering an environment for hosting Jupyter notebooks, and importantly provides access to a large number of open-source data sets and R and Python code.
% This solved a major issue with R programs hosted on e.g. Github, where often the data is not made available.
% Our aim is to run R programs hosted on Kaggle to validate the type signatures we generated for popular R packages.

% \AT{Anything else to say on Kaggle, here? We will probably talk more about the corpus we analyzed later.}
